---
title: "Exploratory Analysis"
output: html_document
---

## Project title

Give your project a concise, interesting title that summarizes the entirety of your project

Analyzing TV Shows and Movies on Netflix Based on Popularity

### Authors

Names of your team members

Yuvika Verma
Jolie Sim
Najat Alnuaimi

### Date

Spring 2024

### Abstract

No more than three sentences that summarize your project. Focus on the very most important aspects. For example: (1) "Our main question is .... This question is important because .... To address the question, we will ...." (2) "We are concerned with ..., because .... To address this concern, we plan to ...." (3) "Consider that .... This is important because .... Accordingly, we plan to ...."

### Keywords

3-5 keywords that summarize your project.
(e.g., "Keywords: human physiology; bicycle exercise; Gen Z; times-series data")

### Introduction  

Briefly introduce your project.  Include 3-5 research questions. What motivates the questions? Why are they important? (at least 200 words)

### Related Work  

Describe your topic and related work in this space. You must include 3 citations to related work (URLs to similar work, high quality articles from the popular press, research papers, etc. ) Please use a standard citation style of your choice. (at least 200 words)

### The Dataset

Where did you find the data? Please include a link to the data source  
Who collected the data?  
How was the data collected or generated?  
Why was the data collected?  
How many observations (rows) are in your data?  
How many features (columns) are in the data?  
What, if any, ethical questions or questions of power do you need to consider when working with this data?  
What are possible limitations or problems with this data?   (at least 200 words)

### Implications

Assuming you answer your research questions, briefly describe the expected or possible implications for technologists, designers, and policymakers. (at least 150 words)

### Limitations & Challenges
What challenges or limitations might you need to address with your project idea more broadly? Briefly discuss. (at least 150 words)

### Summary Information

Write a summary paragraph of findings that includes the 5 values calculated from your summary information R script

### Table

Include a table of aggregate information

Describe why you included the table and what information it reveals

### Chart 1

Include a chart

Describe why you chose this chart and what information it reveals

We chose to include a line graph that displays the amount of content released on Netflix per year because this could show us which years had more content being released compared to others and whether it was around the same time frame as the COVID lockdown/pandemic phase. Through this graph, we can see that there was a significant increase in the amount of content being released on Netflix between 2018-2022, with the peak being in 2019. This goes to show that the amount of content being released on Netflix had seen a growth during the same time as the COVID-19 pandemic (2019-2021). 

Here's an example of how to run an R script inside an RMarkdown file:

``````{r, echo = FALSE, code = readLines("chart1_example.R")}

```


### Chart 2

Include a chart

Describe why you chose this chart and what information it reveals

``````{r, echo = FALSE, code = readLines("line_graph.R")}
```


### Chart 3

Include a chart

Describe why you chose this chart and what information it reveals
